### 2020-09-29 reanalysis.Rmd
bf simulation

---
title: "Re-analysing the data from Moffatt et al. (2020): A textbook illustration of the fallacy of the null hypothesis"

shorttitle: "The fallacy of the null"

author: 
  - name: Ladislas Nalborczyk
    affiliation: 1
    corresponding: yes
    address: GIPSA-lab, CNRS, Univ. Grenoble Alpes, 11 Rue des Mathématiques, 38400 Saint-Martin-d'Hères, France
    email: ladislas.nalborczyk@gipsa-lab.fr
affiliation:
  - id: 1
    institution: Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France

abstract: "Moffatt et al. (2020) reported the results of an experiment (N = 26 in the final sample) comparing the facial (surface) electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on an absence of significant difference in the perioral EMG-amplitude between the rumination and distraction conditions, Moffatt et al. (2020) concluded that (the *self-reported*) inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is hasty and based on waggly evidence. Indeed, concluding on the absence of an effect based on a low-powered statistical test is strongly problematic/uninformative. Moreover, there relation between self-reports and physiological measures... Although the question Moffatt et al. (2020) pursued is of interest, and mehods... Given these limitations, it is unclear whether the target article really adds to the current knowledge. Complete source code, reproducible analyses, and figures are available at https://github.com/lnalborczyk/inner_experience_EMG."

keywords: NHST, Bayesian, fallacy, reanalysis, inner speech, rumination, electromyography

csl: csl/apa7.csl
classoption: man
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false
bibliography: [bib/brms_tutorial.bib, bib/r-references.bib]

nocite: | 
  @R-base, @R-knitr, @R-rmarkdown, @R-tidyverse, @R-papaja, @R-wordcountaddin,
  @R-here

output:
  papaja::apa6_pdf:
    keep_tex: true
    highlight: tango
    latex_engine: xelatex
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(wordcountaddin)
library(ggbeeswarm)
library(rmarkdown)
library(tidyverse)
library(patchwork)
library(bayesplot)
library(ggExtra)
library(papaja)
library(readxl)
library(knitr)
library(BEST)
library(here)
library(glue)

# setting seed for reproducibility
set.seed(666)

# setting up knitr options
knitr::opts_chunk$set(
  cache = TRUE, eval = TRUE, echo = FALSE,
  warning = FALSE, message = FALSE,
  fig.align = "center", dev = "pdf",
  out.width = "100%", fig.pos = "!htb"
  )

# counting words in the current .Rmd file
wordcount <- word_count(knitr::current_input() )
```

Wordcount (excluding abstract, references, tables, and figures): `r wordcount`

\newpage

<!-- NB: You can add comments using these tags -->

# Introduction

...

# Some thoughts on the data

```{r importing_data}
# importing data
df <- read_excel("data/RUM_master_data.xlsx")

# data reshaping (from wide to long)
# NB: 26 participants in the final dataset...

df2 <- df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # extracts muscle name
    mutate(muscle = substr(muscle, 1, 3) ) %>%
    # from long to wide
    pivot_wider(names_from = muscle, values_from = value)
```

```{r babar, fig.cap = "General plot (N = 26)..."}
# position for groups
pd <- position_dodge(0.8)

# plotting the dtaa
df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # mutate(order = factor(order) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.2,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        # draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_dotplot(
        stackdir = "center",
        binaxis = "y",
        position = pd,
        # binwidth = 1,
        dotsize = 0.5,
        alpha = 0.3
        ) +
    stat_summary(
        fun.y = mean,
        geom = "line", size = 1,
        # aes(group = interaction(muscle, order) ),
        aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun.y = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # geom_point(alpha = 0.25, size = 0.5) + 
    # alxtheme +
    facet_grid(~muscle, scales = "free") +
    # scale_color_manual (values = colors2) +
    # scale_fill_manual (values = colors2) +
    labs(x = "Condition", y = "EMG log-amplitude") +
    # theme(legend.title = element_blank() ) +
    # forces lower xlim to 0
    # coord_cartesian(xlim = c(0, 1) ) +
    theme_bw(base_size = 12) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

...

## Concluding on the null from low-powered studies

https://theethicalskeptic.com/2015/08/17/the-four-types-of-null-hypothesis-fallacy/...

```{r}
# importing data
# df <- read_excel("data/RUM_master_data.xlsx")
```

Unfortunately, the same line of reasoning applies for testing the effect of the order, which is even less powered than the test of the main effect of interest, rendering it practically uninformative...

> "In order to test this, a Bayesian paired samples t-test was conducted for the peak log values of muscle activity between the rumination and distraction conditions. This revealed strong evidence in favour of the alternative hypothesis for the FRO muscle (B10 = 18.79), and moderate evidence in favour of the null hypothesis for the OOS (B10 = 0.232) and OOI (B10 = 0.278) muscles, according to current guidelines for interpreting Bayes factors [43]."

...

## Does everyone?

...

```{r everyone, fig.cap = "General plot (mean and 95\\% CI)... showing the inter-individual variability..."}
# does everyone plot
df2 %>%
    pivot_longer(cols = FRO:OOI, names_to = "muscle") %>%
    pivot_wider(names_from = condition, values_from = value) %>%
    mutate(ruminator = ifelse(RUM > DIS, 1, 0) ) %>%
    mutate(ruminator = as.factor(ruminator) ) %>%
    pivot_longer(cols = RUM:BAS, names_to = "condition") %>%
    ggplot(aes(x = condition, y = value, colour = ruminator, fill = ruminator) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.2,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        # draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    # geom_quasirandom(alpha = 0.3) +
    geom_point(size = 2, alpha = 0.8, show.legend = FALSE) +
    geom_line(aes(group = ID), size = 1, alpha = 0.8, show.legend = FALSE) +
    # plitting the by-group mean
    # stat_summary(
    #     fun.y = mean,
    #     geom = "line", size = 2,
    #     aes(group = interaction(muscle, ruminator) ),
    #     position = pd,
    #     show.legend = FALSE
    #     ) +
    # plotting means
    # stat_summary(
    #     fun.y = "mean",
    #     geom = "point", shape = 16, size = 4,
    #     position = pd,
    #     show.legend = TRUE
    #     ) +
    # plotting confidence intervals
    # stat_summary(
    #     fun.data = mean_cl_normal,
    #     geom = "errorbar", size = 2, width = 0,
    #     fun.args = list(mult = 1.96),
    #     show.legend = FALSE,
    #     position = pd
    #     ) +
    # facetting by muscle
    facet_wrap(~muscle) +
    # axis labels
    labs(x = "Condition", y = "EMG log-amplitude") +
    # forces lower xlim to 0
    # coord_cartesian(xlim = c(0, 1) ) +
    theme_bw(base_size = 12) +
    scale_colour_brewer(palette = "Dark2", direction = -1) +
    scale_fill_brewer(palette = "Dark2", direction = -1)
```

Huge inter-individual variability... which leads to the next point, what is the relation between self-reports and EMG?

## Relation between self-report and EMG correlates

...

# Discussion and conclusions

...

# Supplementary materials {#supp}

Reproducible code and figures are available at https://github.com/lnalborczyk/inner_experience_EMG.

# Acknowledgements {-}

...

# References {-}

```{r create_r_references, cache = FALSE}
r_refs(file = "bib/r-references.bib")
```