### 2020-10-06 manuscript.Rmd
adding the dance of the bf

---
title: |
  | **Submitted to Meta-Psychology. Click here to follow the fully transparent editorial process of this submission. Participate in open peer review by commenting through hypothes.is directly on this preprint.**
  | 
  | 
  | Re-analysing the data from @moffatt_inner_2020: A textbook illustration of the absence of evidence fallacy

shorttitle: "The absence of evidence fallacy"

author: 
  - name: Ladislas Nalborczyk
    affiliation: 1
    corresponding: yes
    address: GIPSA-lab, CNRS, Univ. Grenoble Alpes, 11 Rue des Mathématiques, 38400 Saint-Martin-d'Hères, France
    email: ladislas.nalborczyk@gipsa-lab.fr

affiliation:
  - id: 1
    institution: Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France

abstract: "@moffatt_inner_2020 reported the results of an experiment (N = 26 in the final sample) comparing the facial electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on the absence of significant difference in the perioral muscular activity between the rumination and distraction conditions, @moffatt_inner_2020 concluded that *self-reported* inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is at best hasty. Indeed, concluding on the absence of an effect based on an under-powered non-significant *p*-value is strongly problematic/uninformative. Moreover, the relation between self-reports and physiological measures was not directly assessed, but only indirectly inferred from differences (or absence thereof) in group averages. Given the ample inter-individual variability in these measures (as suggested by our reanalysis), we think inferring the individual-level relation between self-reports and physiological measures from group averages is inappropriate. Given these limitations, we conclude that there is limited evidence for the main conclusion put forward by @moffatt_inner_2020 and we suggest ways forward, both from a theoretical and from a methodological perspective. Complete source code, reproducible analyses, and figures are available at https://github.com/lnalborczyk/inner_experience_EMG."

keywords: NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography
csl: csl/apa7.csl
classoption: man, donotrepeattitle
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false

bibliography: [bib/library.bib, bib/r-references.bib]

nocite: | 
  @R-base, @R-knitr, @R-rmarkdown, @R-tidyverse, @R-papaja, @R-wordcountaddin,
  @R-here

output:
  papaja::apa6_pdf:
    keep_tex: true
    highlight: tango
    latex_engine: xelatex

editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(wordcountaddin)
library(BayesFactor)
library(ggbeeswarm)
library(rmarkdown)
library(tidyverse)
library(patchwork)
library(parallel)
library(papaja)
library(readxl)
library(knitr)
library(BEST)
library(brms)
library(glue)
library(here)
library(pwr)

# setting seed for reproducibility
# set.seed(666)

# setting up knitr options
knitr::opts_chunk$set(
  cache = TRUE, eval = TRUE, echo = FALSE,
  warning = FALSE, message = FALSE,
  fig.align = "center", dev = "pdf",
  out.width = "100%", fig.width = 8,
  fig.pos = "!htb"
  )

# counting words in the current .Rmd file
wordcount <- word_count(knitr::current_input() )
```

<!--

NB: You can add comments using these tags

-->

Wordcount (excluding abstract, references, tables, and figures): `r wordcount`

\newpage

# Introduction

The activity of silently talking to oneself or "inner speech" is a foundational ability, allowing oneself to remember, plan, self-motivate or self-regulate [for reviews, see @alderson-day_inner_2015; @perrone-bertolotti_what_2014; @loevenbruck_cognitive_2018]. However, whereas the use inner speech is associated with many adaptive functions in everyday life, inner speech dysfunctions can be identified in multiple psychological disorders. For instance, rumination, broadly defined as unconstructive repetitive thinking about past events and current mood states [@Martin], is involved in the onset and maintenance of serious mental disorders such as depression, anxiety, eating disorders or substance abuse [for a review, see @Nolen-Hoeksema2008].

<!--

Although there are debates about the exact nature of inner speech and whether it is better described as the mere evocation of abstract linguistic representations or as the simulation of executed (overt) speech, there is a considerable amount of evidence that (expanded) inner speech production is accompanied by peripheral muscular activity in the speech effectors [for an overview, see @nalborczyk_understanding_2019]...

-->

Given the predominantly verbal nature of rumination [e.g., @ehring_repetitive_2008; @goldwin_concreteness_2012; @goldwin_concreteness_2013; @mclaughlin_effects_2007], we previously proposed to consider rumination as a form of inner speech and to study it using the methods that have been used to study other forms of inner speech, namely, by using surface electromyography and motor interference protocols [e.g., @nalborczyk_orofacial_2017; @nalborczyk_understanding_2019; @nalborczyk_dissociating_2020; @nalborczyk_articulatory_2020]. We first showed that induced rumination was accompanied by increased facial (both over a forehead and a perioral site) muscular activity in comparison to a rest period [@nalborczyk_orofacial_2017]. However, because rumination was only compared to a rest period, it remained uncertain whether this perioral activity was specifically related to (inner) speech processes. Therefore, we ran an extension of this study, in which we compared verbal to non-verbal rumination, which suggested that the facial EMG correlates we have previously identified were not specifically related to the verbal content of the ruminative thoughts [@nalborczyk_dissociating_2020]. We discussed these findings in length and proposed several theoretical interpretations that can account for these results in the discussion section of @nalborczyk_dissociating_2020 and more extensively in @nalborczyk_understanding_2019. Although these discussions were blatantly ignored by @moffatt_inner_2020, their experimental design nevertheless had the potential to inform our understanding of the involvement of the speech motor system in different varieties of inner speech as well as to clarify the relation between the peripheral correlates of inner speech and the (self-reported) subjective experience.

The main conclusion from @moffatt_inner_2020 is that inner experience between induced rumination and distraction differs "without a change in electromyographic correlates of inner speech". In other words, they suggest that the subjective experience of inner speech is unrelated (or loosely related) to the electromyographic correlates of inner speech, which are thought to be represented mostly by the EMG amplitude recorded over the OOI and OOS muscles. However, for this in-sample observation to be of interest in an out-of-sample context (i.e., to be informative for other non-observed individuals, or said otherwise, to bring information about the population), this absence of difference has to be based on sufficiently powered sample size (given the target effect size) as well as on reliable measures. This is unlikely to be the case here, for reasons that we will present and discuss in the present article. Moreover, a simple visual exploration of the data reveals important variability between individuals in the main effect of interest. That is, some participants had higher perioral (OOS and OOI) muscular activity in the rumination condition than in the distraction condition, and some other participants showed the reverse pattern. This suggests unexplored variation in the determinants of this effect (e.g., the content of the inner experience). Indeed, the relation between the inner experience and the physiological correlates of inner speech production was only inferred from group averages. However, given the important inter-individual variability, this reasoning appears highly problematic. In the following, we explore each of these limitations and suggests ways forward, both from a theoretical and from a methodological perspective.

# Exploring the data

As typical in studies manipulating induced rumination, @moffatt_inner_2020 designed a two-step protocol. First, they aimed to induce a negative mood by asking participants to solve unsolvable or excessively difficult anagram and subtraction tasks. Second, they prompted participants to either ruminate on these (purportedly induced) negative feelings (by asking them to "think about the causes, consequences, and meaning of their current feelings") or to distract themselves (by asking them to "think about a village, city or town that you are particularly familiar with"). Rumination and distraction was manipulated within-subject, with all subjects alternating between rumination and distraction, in a counter-balanced order.

```{r importing_data}
# importing data
df <- read_excel(path = "data/RUM_master_data.xlsx", sheet = 1)

# importing codebook
codebook <- read_excel(path = "data/RUM_master_data.xlsx", sheet = 2)

# data reshaping (from wide to long)
# NB: 26 participants in the final dataset...

# Order in which rumination and distraction inductions were completed
# 1 = rumination induction completed first
# 2 = distraction induction completed first

df2 <- df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # extracts muscle name
    mutate(muscle = substr(muscle, 1, 3) ) %>%
    # from long to wide
    pivot_wider(names_from = muscle, values_from = value)
```

Their final sample of participants, after data exclusion, included 26 participants (data available at https://osf.io/hj7tz/). The EMG data is depicted in Figure \@ref(fig:general) by condition (where `BAS`, `DIS`, and `RUM` refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (`FRO`, `OOI`, `OOS`). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

```{r general, fig.cap = "Average natural logarithm of the EMG peak amplitude per muscle and condition. The black dots and intervals represent the by-group average and 95\\% confidence interval (N = 26). The horizontal white line in the violin plot represents the median. The grey dots represent the individual-level average natural logarithm of the EMG amplitude by muscle and condition."}
# position for groups
pd <- position_dodge(0.8)

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.15,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun = mean,
        geom = "line", size = 1,
        # aes(group = interaction(muscle, order) ),
        aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    # axis labels
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

To model EMG peak amplitude variations in response to the rumination and distraction inductions, we fitted a Bayesian multivariate regression model with the natural logarithm of the EMG peak amplitude as an outcome and *Condition* (baseline, rumination, distraction) as a categorical predictor. Therefore, the intercept represents the estimated natural logarithm of the EMG peak amplitude in the baseline condition, and the slopes for the rumination and distraction conditions represent deviations from the baseline. These analyses were conducted using the `brms` package [@R-brms], an `R` implementation of Bayesian multilevel models that employs the probabilistic programming language `Stan` [@carpenter_stan_2017]. We ran four chains including each 10.000 iterations and a warmup of 2.000 iterations. Posterior convergence was assessed examining autocorrelation and trace plots, as well as the Gelman-Rubin statistic. Constant effects estimates were summarised via their posterior mean and 95% credible interval. We also report Bayes factors (BFs) computed using the Savage-Dickey method.^[This method consists in taking the ratio of the posterior density at the point of interest divided by the prior density at that point [@wagenmakers_bayesian_2010].] These BFs can be interpreted as updating factors, from prior knowledge (what we knew before seeing the data) to posterior knowledge (what we know after seeing the data). A summary of the estimations from this model is presented in Table \@ref(tab:summary). This analysis revealed strong evidence for the hypothesis of a higher average EMG peak amplitude in the rumination condition as compared to the baseline condition for both the FRO and OOI muscles. However, the BFs supported the null hypothesis (i.e., no difference) between the baseline and distraction conditions for the FRO and were inconclusive for both the OOI and OOS muscles.

```{r models, results = "hide"}
##############################
# Bayesian multilevel models #
##############################

# priors for the constant-effects model
priors_constant <- c(
    prior(normal(0, 5), class = b, coef = Intercept, resp = "FRO"),
    prior(normal(0, 1), class = b, resp = "FRO"),
    prior(exponential(1), class = sigma, resp = "FRO"),
    prior(normal(0, 5), class = b, coef = Intercept, resp = "OOS"),
    prior(normal(0, 1), class = b, resp = "OOS"),
    prior(exponential(1), class = sigma, resp = "OOS"),
    prior(normal(0, 5), class = b, coef = Intercept, resp = "OOI"),
    prior(normal(0, 1), class = b, resp = "OOI"),
    prior(exponential(1), class = sigma, resp = "OOI")
    )

# fitting the model
constant_effects <- brm(
    mvbind(FRO, OOS, OOI) ~ 0 + Intercept + condition,
    family = gaussian(),
    prior = priors_constant,
    data = df2,
    chains = 4, cores = detectCores(),
    warmup = 2000, iter = 10000,
    control = list(adapt_delta = 0.95),
    sample_prior = TRUE,
    file = here("models/constant_effects")
    )

# priors for the varying-effects model
# priors_varying <- c(
#     prior(normal(0, 10), class = Intercept, resp = "FRO"),
#     prior(normal(0, 1), class = b, resp = "FRO"),
#     prior(exponential(1), class = sigma, resp = "FRO"),
#     prior(exponential(1), class = sd, resp = "FRO"),
#     prior(normal(0, 10), class = Intercept, resp = "OOS"),
#     prior(normal(0, 1), class = b, resp = "OOS"),
#     prior(exponential(1), class = sigma, resp = "OOS"),
#     prior(exponential(1), class = sd, resp = "OOS"),
#     prior(normal(0, 10), class = Intercept, resp = "OOI"),
#     prior(normal(0, 1), class = b, resp = "OOI"),
#     prior(exponential(1), class = sigma, resp = "OOI"),
#     prior(exponential(1), class = sd, resp = "OOI")
#     )

# fitting the model
# varying_effects <- brm(
#     mvbind(FRO, OOS, OOI) ~ 1 + condition + (1 | ID),
#     family = gaussian(),
#     prior = priors_varying,
#     data = df2,
#     chains = 4, cores = detectCores(),
#     warmup = 2000, iter = 10000,
#     control = list(adapt_delta = 0.95),
#     sample_prior = TRUE,
#     file = here("models/varying_effects")
#     )

# priors for the varying-effects model
priors_varying <- c(
    prior(normal(0, 5), class = b, coef = Intercept, resp = "FRO"),
    prior(normal(0, 1), class = b, resp = "FRO"),
    prior(exponential(1), class = sigma, resp = "FRO"),
    prior(exponential(1), class = sd, resp = "FRO"),
    prior(normal(0, 5), class = b, coef = Intercept, resp = "OOS"),
    prior(normal(0, 1), class = b, resp = "OOS"),
    prior(exponential(1), class = sigma, resp = "OOS"),
    prior(exponential(1), class = sd, resp = "OOS"),
    prior(normal(0, 5), class = b, coef = Intercept, resp = "OOI"),
    prior(normal(0, 1), class = b, resp = "OOI"),
    prior(exponential(1), class = sigma, resp = "OOI"),
    prior(exponential(1), class = sd, resp = "OOI")
    )

# fitting the model (for computing the bf on the intercept)
varying_effects <- brm(
    mvbind(FRO, OOS, OOI) ~ 0 + Intercept + condition + (0 + Intercept | ID),
    family = gaussian(),
    prior = priors_varying,
    data = df2,
    chains = 4, cores = detectCores(),
    warmup = 2000, iter = 10000,
    control = list(adapt_delta = 0.95),
    sample_prior = TRUE,
    file = here("models/varying_effects")
    )
```

```{r bf, eval = FALSE, fig.width = 10, fig.cap = "Savage-Dickey Bayes factor for the difference between the rumination and distraction conditions for each muscle. The BF is computed as the ratio of the posterior density to the prior density at $\\theta = 0$."}
# BF testing
hyp_fro <- hypothesis(
    varying_effects,
    "FRO_conditionDIS = FRO_conditionRUM"
    )

p1 <- plot(hyp_fro, plot = FALSE, theme = theme_bw(base_size = 12) )[[1]] +
    geom_vline(xintercept = 0, linetype = 2) +
    coord_cartesian(xlim = c(-5, 5) ) +
    theme_bw(base_size = 10) +
    theme(legend.title = element_blank() ) +
    labs(
      x = "Log EMG peak amplitude",
      title = paste0(
        "BF10 = ", round(1 / hyp_fro$hypothesis$Evid.Ratio, 3),
        " (BF01 = ", round(hyp_fro$hypothesis$Evid.Ratio, 3), ")"
        )
      )

hyp_ooi <- hypothesis(
    varying_effects,
    "OOI_conditionDIS = OOI_conditionRUM"
    )

p2 <- plot(hyp_ooi, plot = FALSE, theme = theme_bw(base_size = 12) )[[1]] +
    geom_vline(xintercept = 0, linetype = 2) +
    coord_cartesian(xlim = c(-5, 5) ) +
    theme_bw(base_size = 10) +
    theme(legend.title = element_blank() ) +
    labs(
      x = "Log EMG peak amplitude",
      title = paste0(
        "BF10 = ", round(1 / hyp_ooi$hypothesis$Evid.Ratio, 3),
        " (BF01 = ", round(hyp_ooi$hypothesis$Evid.Ratio, 3), ")"
        )
      )

hyp_oos <- hypothesis(
    varying_effects,
    "OOS_conditionDIS = OOS_conditionRUM"
    )

p3 <- plot(hyp_oos, plot = FALSE, theme = theme_bw(base_size = 12) )[[1]] +
    geom_vline(xintercept = 0, linetype = 2) +
    coord_cartesian(xlim = c(-5, 5) ) +
    theme_bw(base_size = 10) +
    theme(legend.title = element_blank() ) +
    labs(
      x = "Log EMG peak amplitude",
      title = paste0(
        "BF10 = ", round(1 / hyp_oos$hypothesis$Evid.Ratio, 3),
        " (BF01 = ", round(hyp_oos$hypothesis$Evid.Ratio, 3), ")"
        )
      )

p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r summary, result = "asis"}
changeSciNot <- function(n) {

  output <- format(n, digits = 4, scientific = TRUE)
  output <- sub("e", "*10^", output)
  output <- sub("\\+0?", "", output)
  output <- sub("-0?", "-", output)
  output

}

summary_varying_effects <-
    # retrieve estimations from bmod5
    summary(varying_effects, prob = 0.95)$fixed %>%
    # convert to dataframe
    data.frame %>%
    # sets rownames as the first column
    rownames_to_column(var = "response") %>%
    # removing ESS information
    dplyr::select(-Bulk_ESS, -Tail_ESS) %>%
    # split response name as two columns
    # separate(col = response, into = c("response", "term"), sep = "_", extra = "merge") %>%
    # rename columns
    magrittr::set_colnames(
        c("Term", "Estimate", "SE", "Lower", "Upper", "Rhat")
        ) %>%
    # compute BF for each effect
    mutate(
        BF10 = 1 / hypothesis(
          varying_effects, glue("{Term} = 0")
          )$hypothesis$Evid.Ratio
        ) %>%
    # round BFs
    # mutate(BF01 = ifelse(BF01 < .001, "<0.001", BF01) ) %>%
    mutate(
      BF10 = ifelse(
        BF10 < 0.001 | BF10 > 1e3,
        changeSciNot(abs(BF10) ),
        round(BF10, 3)
        )
      ) %>%
    # if numeric, round to 3 decimals
    mutate_if(is.numeric, round, 3)

apa_table(
    summary_varying_effects,
    placement = "H",
    digits = 3,
    align = c("l", rep("c", 6) ),
    caption = "Estimated value of the natural logarithm of the EMG peak amplitude in each condition and for each muscle.",
    note = "For each effect, the 'Estimate' reports the estimated average value of the natural logarithm of the EMG peak amplitude, followed by its standard error (SE). The 'Lower' and 'Upper' columns contain the lower and upper bounds of the 95% CrI, whereas the 'Rhat' column reports the Gelman-Rubin statistic. The last column reports the BF in favour of the alternative hypothesis (relative to the null hypothesis).",
    small = TRUE,
    escape = TRUE
    )
```

Because the result of a Bayesian analysis is a joint posterior probability over all parameters of the model, we can compute the posterior distribution of the difference between any pair of conditions. In Figure \@ref(fig:posterior) we represent the posterior distribution of the difference in EMG peak amplitude between the rumination and distraction condition for each muscle. This figure reveals that the most probable value for this difference was $\beta = 0.228$ (95% CrI [0.098, 0.357]) for the FRO muscle, $\beta = 0.081$ (95% CrI [-0.155, 0.324]) for the FRO muscle, and $\beta = 0.047$ (95% CrI [-0.167, 0.27]) for the OOS muscle. Moreover, comparing the posterior distribution to $\theta = 0$ reveals that there is a probability of 0.753 that the average peak EMG amplitude recorded over the OOI is higher in the rumination condition than in the distraction condition [given the model, the priors, and the data from @moffatt_inner_2020].

```{r posterior, fig.width = 10, fig.height = 4, eval = TRUE, fig.cap = "Posterior distribution of the difference in EMG peak amplitude between the rumination and distraction condition for each muscle, along with its mean and 95\\% credible interval."}
# extracts posterior samples
post <- posterior_samples(varying_effects)

# grid of plots
par(mfrow = c(1, 3) )

plotPost(
    post$b_FRO_conditionRUM - post$b_FRO_conditionDIS,
    compVal = 0, xlab = "Rumination - Distraction", main = "FRO"
    )

plotPost(
    post$b_OOI_conditionRUM - post$b_OOI_conditionDIS,
    compVal = 0, xlab = "Rumination - Distraction", main = "OOI"
    )

plotPost(
    post$b_OOS_conditionRUM - post$b_OOS_conditionDIS,
    compVal = 0, xlab = "Rumination - Distraction", main = "OOS"
    )
```

Having nuanced some of the conclusions from @moffatt_inner_2020, we now turn to a discussion of the problems related to conclusions that can be made from under-powered non-significant results.

## Concluding on the null hypothesis from under-powered null-hypothesis significance tests: what could go wrong?

There is an infamous tradition of conducting and interpreting uninformative null-hypothesis significance tests in Psychology [e.g., @harlow_problem_1997; @meehl_appraising_1990-1; @meehl_theoretical_1978; @meehl_theory-testing_1967; @meehl_why_1990]. By "uninformative", we mean that some null-hypothesis significance tests are simply not diagnostic with regards to the substantive effect of interest (e.g., whether there is a difference between conditions A and B).

As highlighted by several authors [e.g., @cohen_earth_1994; @pollard_probability_1987; @rouder_is_2016], concluding that an effect is probably absent solely based on a non-significant *p*-value is the continuous (i.e., probabilistic) extension of the modus tollens and is not a valid argument (i.e., the conclusion does not follow from the premises). This fallacious argument is also known as the *fallacy of acceptance*, the *absence of evidence fallacy* or *the argument from ignorance*, and proceeds as follows: "If the null hypothesis is true, then this observation should *rarely* occur. This observation occurred. Therefore, the null hypothesis is false (or has low probability)". In short, this argument is fallacious because it fails to consider the alternative hypothesis.

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to *severe* tests [e.g., @mayo_severe_2006; @mayo_statistical_2018]. In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives *severe* tests. More precisely, some claim (e.g., $\theta = 0$) is said to be *severely tested* if it had great chances of being falsified, was the claim false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

<!--

More formally, we can define $\text{SEV}(T, x0, H)$, the severity with which claim $H$ passes test $T$ with outcome $x0$, and $\text{SEV}(\mu > \mu_{1}) = \Pr(d(X) \leq d(x0); \mu = \mu_{1})$ the severity of the test the claim $\mu > \mu_{1}$ was submitted to [@mayo_severe_2006; @mayo_statistical_2018]...

-->

Anticipating the legitimate critiques on the power of their study, @moffatt_inner_2020 report the results of a (possibly ran a posteriori) power analysis using the effect size reported in @nalborczyk_orofacial_2017 of $d = 0.72$. This represents a highly optimistic estimate of the substantive effect of interest (i.e., the difference in the natural logarithm of the EMG peak amplitude between the rumination and distraction conditions) as this effect represents the standardised mean difference in EMG amplitude *between a rest and a rumination periods* [@nalborczyk_orofacial_2017].

```{r power, echo = FALSE, eval = FALSE}
# How many participants do we need for a target statistical power of 0.8?
library(pwr)
pwr.t.test(
  d = 0.72, sig.level = 0.05, power = 0.8,
  type = "one.sample", alternative = "two.sided"
  )
```

We suggest the (a priori) power of the study ran by @moffatt_inner_2020 was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter [which seems reasonable given the high inter-individual variability in such effects, cf. the next section but also @nalborczyk_can_2020], therefore the a priori power of the main statistical test from @moffatt_inner_2020 was around $0.44$, meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually $0.36$).

<!--

In other words, the chances of detecting the effect (was the effect actually non-null) were less than obtaining a head in a coin flip.

-->

```{r actual_power, echo = TRUE}
# A priori power for n = 26 and d = 0.36
library(pwr)
pwr.t.test(
  n = 26, d = 0.72 / 2, sig.level = 0.05,
  type = "one.sample", alternative = "two.sided"
  )
```

Once again, anticipating the legitimate critique that the absence of a significant difference is not *necessarily* "significant" evidence for the absence of an effect, @moffatt_inner_2020 reported the following Bayes factor (BF) analysis:

> "[...] therefore it is possible that the sample size of the present study lacked sufficient power to detect the effect of rumination on muscle activity. In order to test this, a Bayesian paired samples t-test was conducted for the peak log values of muscle activity between the rumination and distraction conditions. This revealed strong evidence in favour of the alternative hypothesis for the FRO muscle ($B_{10} = 18.79$), and moderate evidence in favour of the null hypothesis for the OOS ($B_{10} = 0.232$) and OOI ($B_{10} = 0.278$) muscles, according to current guidelines for interpreting Bayes factors [43]."

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the *relative evidence* for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we present below the results of two simulations.

```{r bf-dance, fig.cap = "Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborcayk et al., 2017)."}
# https://julianquandt.com/post/power-analysis-by-data-simulation-in-r-part-ii/
# http://daniellakens.blogspot.com/2016/07/dance-of-bayes-factors.html
# population effect size
cohensd <- 0.36

# sample size
n <- 26

# number of simulated experiments
nsims <- 1e3

# initialises an empty array to store results
results <- data.frame(
  BF10 = numeric(nsims),
  BF10_null = numeric(nsims),
  BF10_double = numeric(nsims)
  )

for (i in 1:nsims) { # for each simulated experiment

  # produce N simulated participants
  x <- rnorm(n = n, mean = 0, sd = 1)

  # produce N simulated participants
  y <- rnorm(n = n, mean = 0 + cohensd, sd = 1)

  # produce N simulated participants
  y_null <- rnorm(n = n, mean = 0, sd = 1)

  # produce N simulated participants
  y_double <- rnorm(n = n, mean = 0 + cohensd * 2, sd = 1)

  results$BF10[i] <- ttestBF(
    x = x, y = y, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  results$BF10_null[i] <- ttestBF(
    x = x, y = y_null, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  results$BF10_double[i] <- ttestBF(
    x = x, y = y_double, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

}

results %>%
  pivot_longer(cols = BF10:BF10_double, names_to = "type") %>%
  mutate(
    type = factor(
      type,
      levels = c("BF10_null", "BF10", "BF10_double"),
      labels = c("Cohen's d = 0", "Cohen's d = 0.36", "Cohen's d = 0.72")
      )
    ) %>%
  ggplot(aes(x = log(value), fill = type) ) +
  geom_density(alpha = 0.5, show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = 2) +
  theme_bw(base_size = 12) +
  facet_wrap(~ type, scales = "free_x") +
  labs(x = "Natural logarithm of the BF10", y = "Density") +
  scale_colour_brewer(palette = "Dark2", direction = 1) +
  scale_fill_brewer(palette = "Dark2", direction = 1)

prop_d0 <- mean(results$BF10_null > 1)
prop_d0.36 <- mean(results$BF10 > 1)
prop_d0.72 <- mean(results$BF10_double > 1)
```

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis), an effect size of d = 0.36 [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of d = 0.72 [i.e., the effect size reported in @nalborczyk_orofacial_2017]. As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis (i.e., when the "true" effect size is d = 0), `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis. When the "true" effect size is of d = 0.36, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of d = 0.72, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by @moffatt_inner_2020. Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis ($\text{BF}_{10}$) using the `BayesFactor` package [@R-BayesFactor]. We used a "medium" prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with $1000$ simulations (i.e., 1000 simulated datasets) for each sample size.

```{r simulated-power, fig.cap = "Average natural logarithm of the Bayes factor in favour of the alternative hypothesis, along with its standard error, computed over 1000 datasets of increasing size simulated from the posterior predictive distribution of the varying-intercept multivariate Bayesian regresion model, fitted on the data from Moffatt et al. (2020). A log-BF belows 0 represents evidence for the null hypothesis (relative to the alternative hypothesis) and a log-BF above 0 represents evidence for the alternative hypothesis (relative to the null hypothesis)."}
# running the simulations
# source("code/bf_simulation.R")

# loading the results
load(file = "results/overall_results.Rda")

# simulations parameters
nsims <- 1e3
sample_size <- seq.int(from = 20, to = 200, by = 10)

# plotting the results
overall_results %>%
    na.omit() %>%
    filter_all(any_vars(. != 0) ) %>%
    pivot_longer(cols = bf_fro:bf_oos, names_to = "bf_type") %>%
    mutate(value = log(value) ) %>%
    mutate(bf_type = factor(bf_type, labels = c("FRO", "OOI", "OOS") ) ) %>%
    group_by(nobs, bf_type) %>%
    summarise(
        across(
            .cols = value,
            .fns = list(mean = mean, se = ~sd(.x) / sqrt(nsims) )
            )
        ) %>%
    ungroup() %>% 
    ggplot(aes(x = nobs, y = value_mean, colour = bf_type, fill = bf_type) ) +
    geom_hline(yintercept = 0, lty = 3) +
    geom_ribbon(
        aes(
          x = nobs, ymin = value_mean - 1.96 * value_se,
          ymax = value_mean + 1.96 * value_se, colour = NULL
          ),
        alpha = 0.5, show.legend = FALSE
        ) + 
    geom_line(show.legend = FALSE) +
    geom_point(show.legend = FALSE) +
    facet_wrap(~bf_type, scales = "free") +
    scale_x_continuous(breaks = unique(sample_size)[c(TRUE, FALSE)]) +
    theme_bw(base_size = 10) +
    labs(x = "Number of participants", y = "Natural logarithm of the Bayes factor") +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

As shown in Figure \@ref(fig:simulated-power), the natural logarithm of the BF in favour of the alternative hypothesis is growing proportionally with sample size. More precisely, whereas BFs computed on small samples (i.e., below 80 participants) support the null hypothesis, BFs computed on larger samples support the alternative hypothesis for all three facial muscles. For instance, the average $\text{BF}_{10}$ computed for the OOI muscle with a sample size of 160 participants is of $\exp(2.18) \approx 8.85$, indicating that these data are approximately $8.85$ times more likely under the alternative hypothesis than under the null hypothesis. To sum up, this reveals that although at low sample sizes, the BF may provide (weak) evidence for the null hypothesis (relative to the alternative hypothesis), this pattern may very well reverse for higher sample sizes.

We should keep in mind some limitations of this analysis, which uses simulated datasets form the posterior predictive distribution estimated on the data collected by @moffatt_inner_2020. This analysis resembles to the Bayesian analogue of the frequentist post-hoc power analysis, which has been much criticised [e.g., @lakens_20_2014]. A crucial assumption of the present analysis is that the data from @moffatt_inner_2020 is our best source of information regarding the main effect of interest. However, the present analysis also differs from the frequentist post-hoc power analysis on several grounds. First, with the present analysis, we do not aim to assess the ability of our statistical test to pass some dichotomic threshold (e.g., accept/reject). Instead, we aim to assess how the $\text{BF}_{10}$ (i.e., the evidence for the alternative hypothesis, relative to the null hypothesis) behaves with varying sample sizes. Second, the present analysis relies on the posterior predictive distribution of the model fitted on the data from @moffatt_inner_2020, which naturally incorporates uncertainty about the effect of interest. By simulating datasets of varying sample sizes from the posterior predictive distribution (and by relying on a large number of simulations), uncertainty about the effect size is naturally incorporated into the results of the simulation.

## Within-subject manipulation of rumination and distraction

In @nalborczyk_dissociating_2020, we manipulated the modality of rumination (whether it is verbal or non-verbal) in a between-subject manner to avoid order effects and to avoid dissipating the effects of the negative mood induction. More precisely, we assumed that inducing rumination after a distraction condition in a within-subject manner would dissipate the effects of the mood induction and therefore reduce the impact of the rumination induction.  In contrast to this approach, @moffatt_inner_2020 asked participants to ruminate and then distract themselves (or reciprocally), after an induced stressor (an induced failure). In Figure \@ref(fig:order), we depict again the EMG data, this time grouped by the order in which the participants went through the rumination and distraction conditions. This figure reveals some potentially interesting differences between the two groups of participants. For instance, the participants that first went through the rumination condition (in green) seem to show a higher increase in the average EMG peak amplitude  recorded over the FRO muscle from baseline than the participants that first went through the distraction condition (in orange).

```{r order, fig.cap = "Average natural logarithm of the EMG peak amplitude by muscle, condition, and group. The green dots and intervals represent the by-group average and 95\\% confidence interval for the participants that first went through the rumination condition, then through the distraction condition. The orange dots and intervals represent the by-group average and 95\\% confidence interval for the participants that first went through the distraction condition, then through the rumination condition. The light green and orange dots in the background represent the individual-level average natural logarithm of the EMG amplitude by muscle, condition, and group."}
# position for groups
pd <- position_dodge(0.8)

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order, labels = c("RUM -> DIS", "DIS -> RUM")) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value, colour = order, fill = order) ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun.y = mean,
        geom = "line", size = 1,
        aes(group = interaction(muscle, order) ),
        # aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun.y = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1) +
    theme(
      legend.title = element_blank(),
      legend.position = "top",
      legend.background = element_rect(size = 0.5, linetype = "solid", colour = "black")
    )
```

Anticipating again that the order of the within-subject conditions may be an issue, @moffatt_inner_2020 say:

> "Unless otherwise reported, the inclusion of order in which the conditions were completed as a between-subjects variable as part of a mixed-design ANOVA produced no significant main effects or interactions involving order."

Unfortunately, the problems we discussed in the next section about the interpretation of under-powered non-significant results also apply to this test. Namely, obtaining a non-significant effect of group is very weak evidence that order did not play a role in the results, given the low power of the tests that were performed. This statistical argument is supported by the brief visual exploration fo the data that we presented in Figure \@ref(fig:order), which suggests possibly crucial differences between the two groups of participants. However, given the sample size in each group (N = 12 and N = 14), it is impossible to know for sure at this point.

## Does everyone show the effect?

We previously noted [e.g., @nalborczyk_orofacial_2017; @nalborczyk_dissociating_2020; @nalborczyk_understanding_2019; @nalborczyk_can_2020] that surface EMG measures of inner speech production were highly variable between individuals. This can be explained by the imagery ability of each individual, the reliability of the measurement, and the instructions that are given to the participants (and whether they are understood in a similar manner by all participants). The data collected by @moffatt_inner_2020 is no exception and present an important degree of inter-individual variability. In Figure \@ref(fig:everyone), we represent again the EMG data for each participant (each line is a participant). We used two colours to represent the participants that showed a higher average EMG peak amplitude either in the rumination condition (in green) or in the distraction condition (in orange). As it can be seen from this figure, whereas some participants show "intermediate" or "ambiguous" (i.e., equivalent) patterns of muscular activity across conditions, some participants show a clear superior EMG peak amplitude in the rumination condition (in green) and some others in the distraction condition (in orange).

```{r everyone, fig.cap = "Inter-individual variability in the main effect of interest (i.e., the difference between the rumination and distraction conditions). Green dots and lines represent the average natural logarithm of the EMG amplitude of participants that showed a higher EMG amplitude in the rumination condition than in the distraction condition, whereas orange dots and lines represent the average natural logarithm of the EMG amplitude of participants that showed a higher EMG amplitude in the distraction condition than in the rumination condition."}
# does everyone plot
df2 %>%
    pivot_longer(cols = FRO:OOI, names_to = "muscle") %>%
    pivot_wider(names_from = condition, values_from = value) %>%
    mutate(ruminator = ifelse(RUM > DIS, 1, 0) ) %>%
    mutate(ruminator = as.factor(ruminator) ) %>%
    mutate(rum_dis = RUM - DIS) %>%
    pivot_longer(cols = RUM:BAS, names_to = "condition") %>%
    ggplot(aes(x = condition, y = value, colour = ruminator) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.15,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_point(size = 2, alpha = 0.8, show.legend = FALSE) +
    # plotting individual trajectories
    geom_line(aes(group = ID), size = 0.5, alpha = 0.8, show.legend = FALSE) +
    # facetting by muscle
    facet_wrap(~muscle) +
    # axis labels
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = -1) +
    scale_fill_brewer(palette = "Dark2", direction = -1)
```

This important inter-individual variability calls into question the use of group averages to describe the nature of inner speech at an individual level. Moreover, this variability suggests that some important confounding factors were not taken into account (i.e., either not manipulated in the experiment or statistically controlled for). In line with @moffatt_inner_2020, we suggest these discrepancies could be explained by differences in the subjective experience of inner speech. We agree that a lot could be learnt by relating this (self-reported) subjective experience to the peripheral muscular correlates of inner speech production. However, this can not be done at the group level, at the risk of missing individual-level patterns. Therefore, we encourage @moffatt_inner_2020 to analyse further their data in order to assess whether the perioral EMG correlates (e.g., the amplitude of the difference between the rumination and distraction conditions on the OOI) can be predicted by the self-reported subjective experience, *at an individual level*.

It should be noted that the question of the qualitative differences in the EMG correlates of inner speech may also be assessed more formally using the model comparison approach developed by @haaf_developing_2017. However, this would require data coming from an experimental design in which inner speech and non-inner speech conditions would be manipulated within-subject and with multiple observations for each participants in each condition [e.g., as in @nalborczyk_can_2020].

<!--

Which leads to the next point, what is the relation between self-reports and EMG?

## Relating the subjective inner experience to the psychophysiological correlates

From page 10: "Independent t-tests revealed that inner speech in the rumination condition was rated as more Condensed if it was engaged in first, compared to if it was completed second (t(29) = 3.237, p = 0.003, d = 1.163). The rumination condition was rated as involving more Positive inner speech by those who engaged in rumination first, compared to those who completed rumina- tion second (t(29) = 2.513, p = 0.018, d = 0.903). In summary, rumination tended to involve more Condensed and Positive inner speech if engaged in first"...

```{r reports, eval = FALSE}
# reshaping from wide to long
EMG <- df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "EMG",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # extracts muscle name
    mutate(muscle = substr(muscle, 1, 3) )

ESM <- df %>%
    select(ID, order = Order, DIS_Future:RUM_Thoughts) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "ESM",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "esm_scale"),
        sep = "_", extra = "merge"
        )

VISQ <- df %>%
    select(ID, order = Order, starts_with("VISQ") ) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with("VISQ"),
        names_to = "visq_item",
        # names_prefix = "wk",
        values_to = "visq_value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = visq_item,
        into = c("visq_item", "condition"),
        sep = "_(?=[^_]+$)"
        # sep = "_", extra = "merge"
        )

# MOOD <- df %>%
#     select(ID, order = Order, PANAS_T1:VAS_Sad_T2) %>%
#     na.omit

VAS <- df %>%
    select(ID, order = Order, VAS_Feelings_RUM:VAS_Problems_DIS) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with("VAS"),
        names_to = "vas_item",
        # names_prefix = "wk",
        values_to = "vas_value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = vas_item,
        into = c("vas_item", "condition"),
        sep = "_(?=[^_]+$)"
        # sep = "_", extra = "merge"
        )

df3 <- bind_cols(c(EMG, ESM, VAS, VISQ) )
```

-->

# Discussion and conclusions

With this paper we aimed to nuance the strong conclusion made by @moffatt_inner_2020, who asserted that the inner experience of rumination was not related to its peripheral muscular correlates. First, we reanalysed the data from @moffatt_inner_2020 and provided some nuance to the conclusion that can be made from these data. Second, we discussed the statistical and epistemological reasons that cast doubt upon the main conclusion of @moffatt_inner_2020. Because the tests conducted by @moffatt_inner_2020 were heavily under-powered, they provide only weak evidence for the absence of difference. Third, we highlighted that the order of the conditions participants went through may impact the effects of the rumination induction (although we can not decide on this issue with the present data). Finally, we showed that the group analyses masked important inter-individual variability that should be more carefully examined.

In addition to these methodological limitations, we now wish to discuss the theoretical interpretations and implications of these results. As discussed in the introduction section, we previously conducted several studies aiming to assess the role of the speech motor system in rumination. Following our initial study [@nalborczyk_orofacial_2017], we ran an extension of this study in which we compared verbal to non-verbal rumination. The results of this study suggested that the facial EMG correlates of verbal and non-verbal rumination were similar [@nalborczyk_dissociating_2020]. Given the ample evidence on the EMG correlates of inner speech production [for an overview, see Chapter 1 in @nalborczyk_understanding_2019], we needed to explain why this particular form of inner speech (induced rumination) was not associated with speech-specific peripheral muscular activity.

In @nalborczyk_dissociating_2020, we suggested that this observation was coherent with the mental-habit view of depressive rumination [@watkins_habit-goal_2014], which defines rumination as a habitual behaviour, automatically triggered by contextual cues such as negative mood. We know habitual behaviours are more automatic (they are not intentionally initiated) than non-habitual behaviours. Interestingly, it has been observed that the automaticity with which a verbal thought is evoked may influence the degree to which it is enacted, that is, the degree to which it recruits the speech motor system [e.g., @cohen_motor_1986; @sokolov_inner_1972]. According to @cohen_motor_1986, the presence of peripheral motor activity during inner speech production may be interpreted in terms of attention sharing. For instance, in novel (hence non-automatic) or difficult situations, the vividness of inner speech may be strengthened by increasing the speech motor activity, resulting in more salient auditory percepts. Relating this idea to the motor control framework we previously proposed [e.g., @loevenbruck_cognitive_2018; @grandchamp_condialint_2019], it may be said that the characteristics of the situation (e.g., novelty, difficulty) may influence the amount of inhibition that is applied to motor commands during inner speech production, hence resulting in more or less visible peripheral muscular activity [see also a discussion of theses ideas in the broader context of motor imagery, @guillot_imagining_2012].

Another possible interpretation is that automatic forms of inner speech may rely more heavily on higher-level (e.g., memory-based) cognitive processes whereas less automatic (i.e., more intentional or deliberate) forms of inner speech may rely more on simulation mechanisms via the use of internal models of the speech motor systems [@nalborczyk_understanding_2019]. In other words, the production of automatic or non-automatic inner speech would be underpinned by different processes that would involve the speech motor system to a different extent. This distinction is similar to the distinction between the two routes of predictions-by-association and prediction-by-simulations in speech perception and comprehension [@pickering_integrated_2013]. The prediction-by-association mechanism would rely more on perceptual sensory experiences and domain-general cognitive abilities whereas the prediction-by-simulation mechanism would rely more simulation of the motor action leading to the speech auditory percept. In the former case, no peripheral muscular activity is expected, whereas in the latter case, the speech motor system would be involved in simulating/emulating the corresponding overt action [cf. also the motor simulation vs. direct simulation (memory retrieval) distinction in @tian_mental_2012]. Whether the physiological correlates of automatic and non-automatic (deliberate) forms of inner speech differ because of inhibitory constraints or because they rely on different processes (e.g., prediction-by-association or prediction-by-simulation) remains an open empirical question. We previously discussed these issues in more length and suggested ways forward from an experimental perspective [cf. the discussion in @nalborczyk_understanding_2019].

To conclude, we wish to bring some nuance to the conclusion of @moffatt_inner_2020, who stated that "In conclusion, induced rumination appeared to involve similar levels of inner speech-related muscle activity to a period of distraction". In consideration of the limitations discussed in the present article, this conclusion seems hasty. Indeed, we provided theoretical (epistemological) and empirical (via simulation) reasons to doubt the strength of the evidence for the null hypothesis in this study. Moreover, supplementary analyses showed that the order of the conditions participants went through may have influenced the effects of the rumination induction on the EMG correlates. Finally, important under-explored inter-individual variability suggests that important determinants of these correlates were not taken into account. We urge the authors to nuance their conclusions, to analyse further their data, and to plan adequately-powered studies in order to settle these issues.

# Supplementary materials {#supp}

Reproducible code and figures are available at https://github.com/lnalborczyk/inner_experience_EMG.

# Acknowledgements {-}

Acknowledgements will be included in the final version of this manuscript.

# References {-}

```{r create_r_references, cache = FALSE}
r_refs(file = "bib/r-references.bib")
```